---
title: "Birthday"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Birthday}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MATH4753F25calebgaines)
```

# Part 1: Mathematical Derivation

We want to compute the probability \( P(A) \) that **at least two** people in a group of size \( n \) share the same birthday.  
Rather than calculating this probability directly, it is easier to calculate the **complementary probability**:

$$
P(A) = 1 - P(A^c),
$$

where \( P(A^c) \) is the probability that no two people share a birthday.

---

### Step 1: Total number of birthday assignments

If there are \( n \) people and \( 365 \) possible birthdays (ignoring leap years), the total number of possible birthday assignments is

$$
365^n.
$$

---

### Step 2: Number of assignments with all distinct birthdays

For the first person, there are 365 possible birthdays.  
For the second person, there are 364 (to avoid matching the first),  
for the third 363, and so on.  

Therefore, the number of ways to assign **distinct birthdays** to all \( n \) people is

$$
365 \times 364 \times 363 \times \cdots \times (365 - n + 1)
= \frac{365!}{(365 - n)!}.
$$

---

### Step 3: Probability that no two people share a birthday

The probability of no shared birthdays is

$$
P(A^c) = 
\frac{\text{# of assignments with all distinct birthdays}}
{\text{total # of assignments}} 
= \frac{\frac{365!}{(365 - n)!}}{365^n}
= \frac{365 \times 364 \times \cdots \times (365 - n + 1)}{365^n}.
$$

---

### Step 4: Probability that at least two people share a birthday

Finally, the desired probability is the complement:

$$
P(A) = 1 - P(A^c) 
= 1 - \frac{365 \times 364 \times \cdots \times (365 - n + 1)}{365^n}.
$$

This gives the exact probability that at least two people in a group of size \( n \) share a birthday.



# Part 2: Reformulation using `lfactorial` and `lchoose`

Start from the original probability expression:
\[
P(A^c)=\frac{365\times 364\times\cdots\times(365-n+1)}{365^n}
=\frac{365!}{(365-n)!\,365^n}.
\]

Rewrite the numerator using the binomial coefficient and factorials:
\[
\frac{365!}{(365-n)!}
= \binom{365}{n} n!.
\]
Thus
\[
P(A^c)=\frac{\binom{365}{n}\,n!}{365^n}.
\]

Take natural logs:
Using \(\log\), \(\log n! = \mathrm{lfactorial}(n)\) and \(\log\binom{365}{n} = \mathrm{lchoose}(365,n)\), we get
\[
\log P(A^c) = \log\big(\binom{365}{n}\big) + \log(n!) - n\log(365)
= \mathrm{lchoose}(365,n) + \mathrm{lfactorial}(n) - n\log(365).
\]

Therefore the probability is
\[
P(A) = 1 - P(A^c)
= 1 - \exp\!\big( \mathrm{lchoose}(365,n) + \mathrm{lfactorial}(n) - n\log(365)\big).
\]



# Part 3: The Function

```{r birthday}
birthday
```



# Part 4: Does it work?

```{r test}
birthday(x = 20:25)
```

## It works!
