---
title: "final-vignette"
name: "Caleb Gaines"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{final-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MATH4753F25calebgaines)
```

# The Package `MATH4753F25calebgaines`

This package was created for the class Applied Statistical Methods (MATH 4753-001) with Professor Wayne Stewart. The package includes 13 functions, 1 test, 2 vignettes, and 1 dataset. This package will hopefully receive a good grade, as it follows the instructions that have been given throughout the class.

# Working Code

## `mynbin.R`

### Description

The function takes values y (total amount of trials), r (trials until success), and p (the probability of success) to calculate P(Y=y)

### Proof it works

```{r mynbin}
y=20
r=7
p=.3
mynbin(y,r,p)
```

### Full Documentation

``` {r doc1}
#' mynbin function
#'
#' @param y total trials until rth success
#' @param r number of successes to be achieved
#' @param p probability of success per trial
#'
#' @returns the probability P(Y=y)
#' @export
#'
#' @examples
#' # single value:
#' mynbin(y=10,r=3,p=.4)
#' # or vector:
#' mynbin(y=3:12,r=3,p=.4)
```

## `mean_sd.R`

### Description

This function returns the mean and standard deviation of a numeric vector x.

### Proof it works

```{r mean_sd}
x <- c(-2.73, 8.15, 0.42, -9.91, 5.56, -0.08, 3.14, -6.89, 7.02, -4.51, 9.27, -1.36, 6.48, -8.05, 1.99, -5.24, 4.76, -3.67, 2.88, -7.40)

mean_sd(x)
```

### Full Documentation

``` {r doc2}
#' Calculate Mean and SD
#'
#' This function returns the mean and standard deviation of a numeric vector.
#'
#' @param x A numeric vector
#' @importFrom stats sd
#' @return A list with mean and sd
#' @export
```

## `myfirstfunction.R`

### Description

This function takes a numeric vector x and returns both a list and a plot of x and y (y=x^2).

### Proof it works

Taking the same x vector as above

```{r myfirstfunction}
myfirstfunction(x)
```

### Full Documentation

```{r doc3}
#' My first function
#'
#' @param x A numeric vector
#'
#' @returns A list with components x and y, where y is the square of x
#' @export
#'
#' @examples
#' myfirstfunction(1:10)
```

## `birthday.R`

### Description

This function takes a vector of values x and calculates the probability for each value that a group of that size shares a birthday.

### Proof it works

```{r birthday}
birthday(18:27)
```

### Full Documentation

``` {r doc4}
#' The Birthday Function
#'
#' @param x a vector of values
#'
#' @returns probability for students to share the same birthday
#' @export
#'
#' @examples birthday(20:24)
```

# A dataset in the package

```{r fire}
fire
```

## The Documentation

```{r firedoc}
#' Fire Damage and Distance Data
#'
#' Data of fire damage and distance from a given file
#'
#' @format ## `fire`
#' A data frame with 15 rows and 2 columns:
#' \describe{
#'   \item{DISTANCE}{Distance from a fire}
#'   \item{DAMAGE}{Damage caused by fire}
#' }
#' @source <FIREDAM.csv from a given file>
```

#### Hopefully this is all good!
